The Symmetry of Partner Modelling

### 2016-02-04 mutual-modeling.tex
Typos

Collaborative learning has often been associated to the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modeling, i.e. that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term 'some
model' to stress the fact that this model is not necessarily details or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner,
upon the modelee's ability to make his state clear to the modeler or upon
the quality of their interactions? We address this question by analyzing
differences between the  accuracy of the models built by different team
members. We therefore report 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that A built about
B was correlated with the accuracy of the model that B built about A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these finding do not
discard the fact individual abilities also  participate in the quality of
modellng process .

@@

In summary, the controversy around the cognitive depth of shared understanding
pertains to psycholinguistic, which investigates 'natural conversations, but  in
collaborative learning, the situation is different; tasks are precisely designed
for requirint a deeper negotiation fo meaning.  The craft of CSCL is to design
collaborative situations that create interdependence, avoid groupthink, and
allows students to detect any 'illusion of shared understanding' (Cherubini et
al, 2005). Our question is hence not  anymore"do peers build a shared
understanding ?", but becomes "when peers have to build a shared understanding,
how do they do it ?".  This question addresses the mechanisms of grounding, in
which the basic sequence is: make a proposition, detect misunderstanding and, if
any, repair them.  This paper focuses on the middle part, the detection of
misunderstandings.  Detecting a misunderstanding means that the emitter of a
message identifies a mismatch between his communicative intention and the way
his message is understood by his or her partner. Detecting one peer's
understanding is investigate hereafter under the general umbrella of partner
modelling.

@@

To answer this question, we went back to five previous studies that addressed
various other research questions, but in which the correlation between the
quality of these two models could be computed.  The studies we report do hence
not constitue a clean sequence of experiment to investigate mutual modelling but
the ad-hoc revisiting of previous experiments to explore a question that we had
then neglected. Some of these studies are about collaborative learning while
others are only collaborative problem solving, but the latter rely on rich tasks
that are similar to those we use in CSCL.

@@

Our method for investigating mutual modelling is observe the variations of
accuracy that result from variations of external parameters that appear as
variable in the formula presented in the previous section: the accuracy should
for go down if we increase the cost of grounding acts and goes up if we increase
the grounding criterion (the necessity to build a shared understanding). One way
to produce these variations is what CSCW,researchers have called ``awareness
tools'', \ie functionalities that inform $A$ is about $B$'s actions that A could
not directly perceive because $B$ was working on a different subset of the
virtual space. Different awareness tools are used in the following studies as
methodological tricks  to manipulate experimentally the  mutual modelling
activity.

@@

We have however to bring some nuances to the social viewpoint
($\mathcal{H}_{3}$).  The main feature that can be associated to the team level
in this experiment is probably not the quality of their verbal interactions per
se (they interact mostly by drawing on a shared map), but rather the consistency
of  the spatial exploration strategy: a clear strategy facilitates memorizing
one's partner path.  One could argue whether the team strategy can be
dissociated from the team interactions quality or constitues one of its
components.

@@

\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
between self and social accuracy here because we do not have an estimation of
self-accuracy: subjects describe their emotions but we have no way to check if
there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
interpret this too far, but to show that there are large variations even within one
area  (perceiving emotions). These variations shomehow question the existence of
a general aptitude to model others ( $\mathcal{H}_{1}$).

@@

We acknowledge that the difference between these hypotheses is  rather
theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$)
and the process of helping one's partner to model onself ($\mathcal{H}_{3}$) are
mediated by verbal interactions in the team. One could hardly imagine someone
managing a very accurate modelling despite the very bad quality of interactions.
There is a bidirectional causal link between the accuracy of mutual modelling
and the quality of interactions. This rather artificial distinction does however
participate to the more fundamental discussion of the role of individual and
social mechanisms in human cognition. In the field of social cognition, it is a
common place to state that 'the whole is greater than the sum of the parts'.
This refers to the emergence of team properties than cannot be reduced to the
set of individual contributions.  Our paper illustrates this emergence by
showing the symmetry of mutual modelling

>>>>

Collaborative learning has often been associated to the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modelling, i.e. that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term 'some
model' to stress the fact that this model is not necessarily details or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner,
upon the modelee's ability to make his state clear to the modeler or upon
the quality of their interactions? We address this question by analyzing
differences between the  accuracy of the models built by different team
members. We therefore report 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that A built about
B was correlated with the accuracy of the model that B built about A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these finding do not
discard the fact individual abilities also  participate in the quality of
modelling process .

@@

In summary, the controversy around the cognitive depth of shared understanding
pertains to psycholinguistic, which investigates 'natural conversations, but  in
collaborative learning, the situation is different; tasks are precisely designed
for requiring a deeper negotiation of meaning.  The craft of CSCL is to design
collaborative situations that create interdependence, avoid group think, and
allows students to detect any 'illusion of shared understanding' (Cherubini et
al, 2005). Our question is hence not  anymore"do peers build a shared
understanding ?", but becomes "when peers have to build a shared understanding,
how do they do it ?".  This question addresses the mechanisms of grounding, in
which the basic sequence is: make a proposition, detect misunderstanding and, if
any, repair them.  This paper focuses on the middle part, the detection of
misunderstandings.  Detecting a misunderstanding means that the emitter of a
message identifies a mismatch between his communicative intention and the way
his message is understood by his or her partner. Detecting one peer's
understanding is investigate hereafter under the general umbrella of partner
modelling.

@@

To answer this question, we went back to five previous studies that addressed
various other research questions, but in which the correlation between the
quality of these two models could be computed.  The studies we report do hence
not constitute a clean sequence of experiment to investigate mutual modelling but
the ad-hoc revisiting of previous experiments to explore a question that we had
then neglected. Some of these studies are about collaborative learning while
others are only collaborative problem solving, but the latter rely on rich tasks
that are similar to those we use in CSCL.

@@

Our method for investigating mutual modelling is observe the variations of
accuracy that result from variations of external parameters that appear as
variable in the formula presented in the previous section: the accuracy should
for go down if we increase the cost of grounding acts and goes up if we increase
the grounding criterion (the necessity to build a shared understanding). One way
to produce these variations is what CSCW researchers have called ``awareness
tools'', \ie functionalities that inform $A$ is about $B$'s actions that A could
not directly perceive because $B$ was working on a different subset of the
virtual space. Different awareness tools are used in the following studies as
methodological tricks  to manipulate experimentally the  mutual modelling
activity.

@@

We have however to bring some nuances to the social viewpoint
($\mathcal{H}_{3}$).  The main feature that can be associated to the team level
in this experiment is probably not the quality of their verbal interactions per
se (they interact mostly by drawing on a shared map), but rather the consistency
of  the spatial exploration strategy: a clear strategy facilitates memorizing
one's partner path.  One could argue whether the team strategy can be
dissociated from the team interactions quality or constitutes one of its
components.

@@

\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
between self and social accuracy here because we do not have an estimation of
self-accuracy: subjects describe their emotions but we have no way to check if
there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
interpret this too far, but to show that there are large variations even within one
area  (perceiving emotions). These variations somehow question the existence of
a general aptitude to model others ( $\mathcal{H}_{1}$).

@@

We acknowledge that the difference between these hypotheses is  rather
theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$)
and the process of helping one's partner to model oneself ($\mathcal{H}_{3}$) are
mediated by verbal interactions in the team. One could hardly imagine someone
managing a very accurate modelling despite the very bad quality of interactions.
There is a bidirectional causal link between the accuracy of mutual modelling
and the quality of interactions. This rather artificial distinction does however
participate to the more fundamental discussion of the role of individual and
social mechanisms in human cognition. In the field of social cognition, it is a
common place to state that 'the whole is greater than the sum of the parts'.
This refers to the emergence of team properties than cannot be reduced to the
set of individual contributions.  Our paper illustrates this emergence by
showing the symmetry of mutual modelling
